Static and Dynamic Bethe-Salpeter Equations in the $T$-Matrix Approximation

### 2022-02-14 Manuscript/Tmatrix.tex
OK for T2

Based on Rayleigh-Schr\"odinger perturbation theory, the non-linear eigenproblem \eqref{eq:dBSE} can be split as a zeroth-order static reference and a first-order dynamic perturbation, such that
\begin{multline}
\label{eq:LR-PT}
	\begin{pmatrix}
		\bA{}{\dBSE}(\Om{S}{})		&	\bB{}{\BSE}(\Om{S}{})	\\
		-\bB{}{\BSE}(-\Om{S}{})	&	-\bA{}{\BSE}(-\Om{S}{})	\\
	\end{pmatrix}
	\\
	=
	\begin{pmatrix}
		\bA{}{\BSE}	&	\bB{}{\BSE}	
		\\
		-\bB{}{\BSE}	&	-\bA{}{\BSE}	
		\\
	\end{pmatrix}
	+
	\begin{pmatrix}
		\bA{}{(1)}(\Om{S}{})		&	\bO	\\
		\bO	&	-\bA{}{(1)}(-\Om{S}{})	\\
	\end{pmatrix}
\end{multline}
with
\begin{equation}
	\label{eq:BSE-A1}
	A_{ia,jb}^{(1)}(\omega) = \widetilde{T}_{ib,aj}(\omega) - T_{ib,aj}(\omega = 0)
\end{equation}
As usual, one can naturally expand the $S$th BSE excitation energy and its corresponding eigenvector as
\begin{subequations}
\begin{gather}
	\Om{S}{} = \Om{S}{\BSE} + \Om{S}{(1)} + \ldots,
	\\
	\begin{pmatrix}
		\bX{S}{}	\\
		\bY{S}{}	\\
	\end{pmatrix}
	= 
	\begin{pmatrix}
		\bX{S}{\BSE}	\\
		\bY{S}{\BSE}	\\
	\end{pmatrix}
	+
	\begin{pmatrix}
		\bX{S}{(1)}	\\
		\bY{S}{(1)}	\\
	\end{pmatrix}
	+ \ldots
\end{gather}
\end{subequations}
Solving the static BSE [see Eq.~\eqref{eq:BSE}] yields the (zeroth-order) static $\Om{S}{\BSE}$ excitation energies and their corresponding eigenvectors $\bX{S}{\BSE}$ and $\bY{S}{\BSE}$. 
The first-order correction to the $S$th excitation energy is, within the dTDA,
\begin{equation}
\label{eq:Om1-TDA}
	\Om{S}{(1)} = \T{(\bX{S}{\BSE})} \cdot \bA{}{(1)}(\Om{S}{\BSE}) \cdot \bX{S}{\BSE}
\end{equation}
This correction can be renormalized by computing, at no extra cost, the renormalization factor which reads
\begin{equation}
\label{eq:Z}
	\zeta_{S} = \qty[ 1 - \T{(\bX{S}{\BSE})} \cdot \left. \pdv{\bA{}{(1)}(\Om{S}{})}{\Om{S}{}} \right|_{\Om{S}{} = \Om{S}{\BSE}} \cdot \bX{S}{\BSE} ]^{-1}
\end{equation}
This yields our final expression for the dynamically-corrected BSE excitation energies:
\begin{equation}
	\Om{S}{\text{dyn}} = \Om{S}{\text{stat}} + \Delta\Om{S}{\text{dyn}} = \Om{S}{\BSE} + \zeta_{S} \Om{S}{(1)}
\end{equation}
Note again that the present perturbative scheme does not allow to access double excitations as only excitations calculated within the static approach can be dynamically corrected.

@@

In terms of computational cost, the overall scaling of BSE@$GT$ is equivalent to BSE@$GW$ as they both correspond to seeking the lowest eigenvalues of a matrix of size $(2\Nocc\Nvir \times 2\Nocc\Nvir)$. 
Searching iteratively for the lowest eigenstates can be routinely performed via Davidson's algorithm with a $\order*{\Norb^4}$ computational cost. \cite{Davidson_1975}
The cost of the dynamical correction, which is thoroughly discussed in Ref.~\onlinecite{Loos_2020h}, is more expensive but is again equivalent in both formalisms.
The computational cost associated with the computation of the $T$-matrix and the screening $W$ both scale as $\order*{\Norb^6}$ as one must obtain, respectively, all the eigenvalues and eigenvectors of the pp-RPA and ph-RPA problems.
However, the prefactor of the pp-RPA calculation is significantly larger than its ph-RPA counterpart due to the larger size of the pp-RPA matrices and its non-Hermitian nature. \cite{Peng_2013,Scuseria_2013,Yang_2013,Yang_2013b,Yang_2014a,Zhang_2015,Zhang_2016}
Moreover, within the $T$-matrix formalism, one must compute both the singlet and triplet contributions of the $T$-matrix, while for singlet states, only the singlet part of $W$ is required.
Although similar approaches remain to be developed for the $T$-matrix formalism, contour deformation and density fitting techniques can be efficiently implemented in the case of $GW$ to reduce the scaling to $\order*{\Norb^3}$. \cite{Duchemin_2019,Duchemin_2020,Duchemin_2021}

@@

As a second example, we consider the symmetric dissociation of the linear molecule beryllium hydride (\ce{BeH2}), a system for which one can assume that the screening plays a more important role than in the previous example.
The variation of the lowest singlet and triplet excitation energies as a function of the distance $R_{\ce{Be-H}}$ is shown in Fig.~\ref{fig:BeH2}, while the quasiparticle energies of the frontier orbitals and the associated (fundamental) gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:BeH2_gap}.
Again, one notes that the $G_0W_0$ and $G_0T_0$ quasiparticle energies are very similar near the Fermi level.
Therefore, one can safely assume that any significant variation of the excitation energies computed within the $GW$- and $GT$-based formalisms originates mainly from their distinct kernel.
The excitation energies computed with the dynamical schemes, dBSE@$G_0W_0$ and dBSE@$G_0T_0$, are reported as thin solid lines.
Here, one can show that dynamical corrections improves in most cases the agreement between BSE and FCI.
All calculations are performed with the cc-pVDZ basis.

For the four lowest singlet excited states (left panel of Fig.~\ref{fig:BeH2}), dBSE@$G_0T_0$ is clearly better than dBSE@$G_0W_0$, while the opposite trend is observed for the four lowest triplet states (right panel of Fig.~\ref{fig:BeH2}).
Note that, for large $R_{\ce{Be-H}}$, the two BSE-based schemes provide only a qualitative description of the excited states with error of several \si{\eV}.
Nonetheless, the overall ordering of the excited states are globally respected.

@@

%%% FIGURE 6 %%%
\begin{figure*}
	\includegraphics[width=\linewidth]{H2O}
	\caption{Error with respect to FCI for the singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{H2O} at equilibrium geometry computed at various levels of theory with the aug-cc-pVDZ basis.
	The geometry as well as the reference FCI values have been extracted from the QUEST database. \cite{Veril_2021}
	Raw data are reported in {\SupInf}.}
	\label{fig:H2O}
\end{figure*}
%%% %%% %%% %%%

As a third and final example, we compute the excitation energies associated with the two lowest singlet and two lowest triplet excited states of water at equilibrium geometry (see Fig.~\ref{fig:H2O}).
Note that all these excited states are of Rydberg nature and correspond to $n \to 3s$ and $n \to 3p$ transitions for the $B_1$ and $A_2$ states, respectively. \cite{Loos_2018a}
In addition to the BSE-based models studied in the present manuscript, we have selected well-known wave function methods, \cite{Bene_1971,Head-Gordon_1994,Head-Gordon_1995,Dreuw_2005} namely CIS, CIS(D), TDHF, and FCI (taken as reference) and computed the excitation energies for these transitions.
Two key observations can be made:
i) BSE@$G_0W_0$ is by far the best performer with a slight overestimation of the order of \SI{0.1}{eV} (as compared to FCI);
ii) BSE@$G_0T_0$ systematically underestimates the excitation energies [similarly to CIS(D)] and outperforms CIS, CIS(D) and TDHF for the singlet states only. 
These general trends are also observed for other systems and they nicely evidence the crucial role of the screening in $GW$, hinting that a screened version of the $T$-matrix formalism as proposed in Ref.~\onlinecite{Romaniello_2012} might be a promising way for improvement.

@@

In order to derive the dynamical kernel $\tilde{T}_{ib,aj}$ given in Eq.~\eqref{eq:dT}, we follow Ref.~\onlinecite{Strinati_1980} (see also Ref.~\onlinecite{Loos_2020e}) and start from the equation for the BSE  amplitude
\begin{equation}
\label{eq:Strinati_1}
	\chi_S(1,1') = \int d3d4d5d6L_0(1,4,1',3)\Xi(3,5,4,6) \chi_S(6,5)
\end{equation}
where $L_0$ is given by Eq.~\eqref{eq:L0} and the $T$-matrix kernel is
\begin{equation}
\label{eq:T-matrix_kernel}
	\Xi(3,5,4,6) =  i\frac{\delta \Sigma(3,4)}{\delta G(6,5)}\approx -T(3,5,4,6)
\end{equation}
Equation (\ref{eq:Strinati_1}) is derived by assuming that i) the (resonant) pole $\omega_S=E_S-E_0>0$ of $L$ is isolated from the other poles (which is usually the case for neutral excitations in finite systems), and  ii) the poles of $L_0$ are different form $\omega_S$ (which is also generally the case).
The so-called $T$-matrix self-energy $\Sigma$ is given by Eq.~\eqref{eq:SigGT} with \cite{Martin_2016,Romaniello_2012}
\begin{equation}
\label{eq:T-matrix}
	T(3,8,4,7) =
	- v(3,8)\delta(3,4)\delta(7,8)
	+ v(3,8)\delta(4,8)\delta(3,7)
	+ i \int d1'd2' v(3,8)G(3,1')G(8,2')T(1',2',4,7)
\end{equation}  
where $v$ is the bare Coulomb operator and we neglect the functional derivative $\delta T/\delta G$ in the kernel $\Xi$. 
The first two terms in the right-hand side of Eq.~\eqref{eq:T-matrix} are the Hartree and exchange contributions to the $T$-matrix, whereas the last term is the correlation contribution.
Making the time dependence of Eq.~\eqref{eq:Strinati_1} explicit and defining $T(3,5,4,6)=- \delta(\tau^+_{35})\delta(\tau^+_{64})\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})$, one gets 
\begin{equation}
\label{eq:DYN_T}
	\chi_S(\bx_1,\bx_{1'},\tau_{11'})e^{-i\omega_S(t_1+t_{1'})/2}
	=-i\int d3d4d5d6 G(\bx_1,\bx_3;\tau_{13})G(\bx_4,\bx_{1'};\tau_{11'})
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}
\end{equation}
[where $\tau_{ij}=t_i-t_j$ and $\tau^+_{ij}=t^+_i-t_j$ with $t^+_i=t_i+\eta$ ($\eta \to 0^+$)] and
\begin{multline}
\label{eq:T_bar}
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34}) 
	= v(\bx_3,\bx_5)\delta(\tau_{34})\delta(\bx_3,\bx_4)\delta(\bx_6,\bx_5)-v(\bx_3,\bx_5)\delta(\tau_{34})	\delta(\bx_3,\bx_6)\delta(\bx_4,\bx_5)
	\\
	+ i \int d7 d8 v(\bx_3,\bx_5)G(\bx_3,\bx_7;\tau_{38})G(\bx_5,\bx_8;\tau^+_{38})\Bar{T}(\bx_7,\bx_8,\bx_4,\bx_6;\tau_{84}).
\end{multline}
Using the Fourier transform $G(\tau)=\int \frac{d\omega}{2\pi}G(\omega)e^{-i\omega\tau}$, changing variable from $t_3$ to $\tau_{34}$ and taking the limit $t_{1'}=t_1^+$, we have
\begin{multline}
	\chi_S(\bx_1,\bx_{1'},0^-)
	= -i\int d\bx_3 d\bx_4 d\bx_5 d\bx_6\int d\tau_{34} \int \frac{d\omega'}{2\pi}G(\bx_1,\bx_3;\omega'+\omega_S)G(\bx_4,\bx_{1'};\omega')e^{i\omega'\tau_{34}}
	\\
	 \times \Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}.
\end{multline}
Using the Lehman representation of the one-body Green's function in the quasiparticle approximation given by Eq.~\eqref{eq:G},
%$$G(x_1,x_2;\omega)=\sum_{p}\frac{\psi_p(x_1)\psi_p^*(x_2)}{\omega-\epsilon_p+ i\eta \text{sgn}(\epsilon_p-\mu)},$$ where $\psi_p$ and $\epsilon_p$ are single-particle wavefunctions and energies, respectively, and $\mu$ the chemical potential, 
and multiply the left- and right-hand sides by $(\e{a}{}-\e{i}{}-\omega_S)\int d\bx_1 d\bx^{\prime}_1 \SO{a}(\bx_1)\SO{i}(\bx_1')$, we obtain  
\begin{multline}
\label{eq:General_DYN}
	\qty(\e{a}{}-\e{i}{}-\omega_S)\int d\bx_1d\bx^{\prime}_1 \SO{a}(\bx_1)\SO{i}(\bx_1')  \chi_S(\bx_1,\bx_{1'},0^-) =
	-\int d\bx_3 d\bx_4 d\bx_5 d\bx_6\int d\tau_{34}\SO{a}(\bx_3)\SO{i}(\bx_4)
	\\
	\times \qty[\Theta(\tau_{34})e^{i\e{i}{}\tau_{34}}+\Theta(-\tau_{34})e^{i(\e{a}{}-\omega_S)\tau_{34}} ]
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{i\omega_S\tau_{34}/2}
\end{multline}
(where $\Theta$ is the Heaviside step function) using the fact that
\begin{equation}
	\Theta(\pm\tau)e^{-i\alpha \tau}=\mp\frac{1}{2\pi i}\lim_{\eta\to 0^+}\int d\omega \frac{1}{\omega-\alpha\pm i\eta}e^{-i\omega\tau}
\end{equation}

@@

\begin{equation}
	\label{eq:egval}
	\qty(\e{a}{}-\e{i}{}-\omega_S) X_{ia,S}+\sum_{jb}X_{jb,S}\mel{ib}{}{aj} - \sum_{jb}X_{jb,S}\widetilde{T}_{ib,aj}(\omega_S)=0
\end{equation}
%with $$\langle ab|ij\rangle=\int d\bx_1d\bx_2 \SO{a}(\bx_1) \SO{b}(\bx_2)v(\bx_1,\bx_2) \SO{i}(\bx_1) \SO{j}(\bx_2),$$ 
where the spectral representation of the dynamical $T$-matrix is
\begin{equation}
	\widetilde{T}_{ib,aj}(\omega_S)
	= \frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\Bar{T}^c_{ib,aj}(\omega)e^{-i\omega\eta} \left[\frac{1}{\omega_S-\omega+\e{j}{}+\e{i}{}+i\eta}+\frac{1}{\omega_S+\omega-\e{b}{}-\e{a}{}+i\eta}\right]. 
 \label{eq:H2preso_eigen_2_T}
\end{equation}
Equation \eqref{eq:egval} represents a non-linear eigenvalue equation to calculate the positive excitation energies of a system, which can be rewritten as
\begin{equation}
\label{eq:H2preso_eigen}
	\sum_{jb} A_{ia,jb}(\omega_S) X_{jb,S}=\omega_S X_{ia,S}
\end{equation}
with
\begin{equation}
	A_{ia,jb}(\omega_S)=(\e{a}{}-\e{i}{})\delta_{ij}\delta_{ab} + \mel{ib}{}{aj} - \widetilde{T}_{ib,aj}(\omega_S).
\end{equation}
%Here $H_{ia,jb}(\omega_S)=A_{ia,jb}(\omega_S)$ and $H_{ia,bj}(\omega_S)=B_{ia,jb}(\omega_S)$.
If one drops the dynamical part $\widetilde{T}$, one ends up with the usual time-dependent Hartree-Fock (TDHF) equations. \cite{Dreuw_2005}
To calculate the correlation contribution, one must employ Eq.~\eqref{eq:T} in Eq.~\eqref{eq:H2preso_eigen_2_T}, and after integration over the frequency, it yields Eq.~\eqref{eq:dT}.

>>>>

Based on Rayleigh-Schr\"odinger perturbation theory, the non-linear eigenproblem \eqref{eq:dBSE} can be split as a zeroth-order static reference and a first-order dynamic perturbation, such that
\begin{multline}
\label{eq:LR-PT}
	\begin{pmatrix}
		\bA{}{\dBSE}(\Om{S}{})		&	\bB{}{\BSE}(\Om{S}{})	\\
		-\bB{}{\BSE}(-\Om{S}{})	&	-\bA{}{\BSE}(-\Om{S}{})	\\
	\end{pmatrix}
	\\
	=
	\begin{pmatrix}
		\bA{}{\BSE}	&	\bB{}{\BSE}	
		\\
		-\bB{}{\BSE}	&	-\bA{}{\BSE}	
		\\
	\end{pmatrix}
	+
	\begin{pmatrix}
		\bA{}{(1)}(\Om{S}{})		&	\bO	\\
		\bO	&	-\bA{}{(1)}(-\Om{S}{})	\\
	\end{pmatrix}
\end{multline}
with
\begin{equation}
	\label{eq:BSE-A1}
	A_{ia,jb}^{(1)}(\omega) = \widetilde{T}_{ib,aj}(\omega) - T_{ib,aj}(\omega = 0)
\end{equation}
As usual, one can naturally expand the $S$th BSE excitation energy and its corresponding eigenvector as
\begin{subequations}
\begin{gather}
	\Om{S}{} = \Om{S}{\BSE} + \Om{S}{(1)} + \ldots,
	\\
	\begin{pmatrix}
		\bX{S}{}	\\
		\bY{S}{}	\\
	\end{pmatrix}
	= 
	\begin{pmatrix}
		\bX{S}{\BSE}	\\
		\bY{S}{\BSE}	\\
	\end{pmatrix}
	+
	\begin{pmatrix}
		\bX{S}{(1)}	\\
		\bY{S}{(1)}	\\
	\end{pmatrix}
	+ \ldots
\end{gather}
\end{subequations}
Solving the static BSE [see Eq.~\eqref{eq:BSE}] yields the (zeroth-order) static $\Om{S}{\BSE}$ excitation energies and their corresponding eigenvectors $\bX{S}{\BSE}$ and $\bY{S}{\BSE}$. 
The first-order correction to the $S$th excitation energy is, within the dTDA,
\begin{equation}
\label{eq:Om1-TDA}
	\Om{S}{(1)} = \T{(\bX{S}{\BSE})} \cdot \bA{}{(1)}(\Om{S}{\BSE}) \cdot \bX{S}{\BSE}
\end{equation}
This correction can be renormalized by computing, at no extra cost, the renormalization factor which reads
\begin{equation}
\label{eq:Z}
	\zeta_{S} = \qty[ 1 - \T{(\bX{S}{\BSE})} \cdot \eval{\pdv{\bA{}{(1)}(\Om{S}{})}{\Om{S}{}}}_{\Om{S}{} = \Om{S}{\BSE}} \cdot \bX{S}{\BSE} ]^{-1}
\end{equation}
This yields our final expression for the dynamically-corrected BSE excitation energies:
\begin{equation}
	\Om{S}{\text{dyn}} = \Om{S}{\text{stat}} + \Delta\Om{S}{\text{dyn}} = \Om{S}{\BSE} + \zeta_{S} \Om{S}{(1)}
\end{equation}
Note again that the present perturbative scheme does not allow to access double excitations as only excitations calculated within the static approach can be dynamically corrected.

@@

In terms of computational cost, the overall scaling of BSE@$GT$ is equivalent to BSE@$GW$ as they both correspond to seeking the lowest eigenvalues of a matrix of size $(2\Nocc\Nvir \times 2\Nocc\Nvir)$. 
Searching iteratively for the lowest eigenstates can be routinely performed via Davidson's algorithm with a $\order*{\Norb^4}$ computational cost. \cite{Davidson_1975}
The cost of the dynamical correction, which is thoroughly discussed in Ref.~\onlinecite{Loos_2020h}, is more expensive but is again equivalent in both formalisms.
The computational cost associated with the computation of the $T$-matrix and the screening $W$ both scale as $\order*{\Norb^6}$ as one must obtain all the eigenvalues and eigenvectors of the pp-RPA and ph-RPA problems, respectively.
However, the prefactor of the pp-RPA calculation is significantly larger than its ph-RPA counterpart due to the larger size of the pp-RPA matrices and its non-Hermitian nature. \cite{Peng_2013,Scuseria_2013,Yang_2013,Yang_2013b,Yang_2014a,Zhang_2015,Zhang_2016}
Moreover, within the $T$-matrix formalism, one must compute both the singlet and triplet contributions of the $T$-matrix, while for singlet states, only the singlet part of $W$ is required.
Although similar approaches remain to be developed for the $T$-matrix formalism, contour deformation and density fitting techniques can be efficiently implemented in the case of $GW$ to reduce the scaling to $\order*{\Norb^3}$. \cite{Duchemin_2019,Duchemin_2020,Duchemin_2021}

@@

As a second example, we consider the symmetric dissociation of the linear molecule beryllium hydride (\ce{BeH2}), a system for which one can assume that the screening plays a more important role than in the previous example.
The variation of the lowest singlet and triplet excitation energies as a function of the distance $R_{\ce{Be-H}}$ is shown in Fig.~\ref{fig:BeH2}, while the quasiparticle energies of the frontier orbitals and the associated (fundamental) gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:BeH2_gap}.
All calculations are performed with the cc-pVDZ basis.
Again, one notes that the $G_0W_0$ and $G_0T_0$ quasiparticle energies are very similar near the Fermi level.
Therefore, one can safely assume that any significant variation of the excitation energies computed within the $GW$- and $GT$-based formalisms originates mainly from their distinct kernel.
The excitation energies computed with the dynamical schemes, dBSE@$G_0W_0$ and dBSE@$G_0T_0$, are reported as thin solid lines.
Here, one can show that dynamical corrections improves in most cases the agreement between BSE and FCI.

For the four lowest singlet excited states (left panel of Fig.~\ref{fig:BeH2}), dBSE@$G_0T_0$ is clearly better than dBSE@$G_0W_0$, while the opposite trend is observed for the four lowest triplet states (right panel of Fig.~\ref{fig:BeH2}).
Note that, for large $R_{\ce{Be-H}}$, the two BSE-based schemes provide only a qualitative description of the excited states with errors of several \si{\eV}.
Nonetheless, the overall ordering of the excited states are globally respected.

@@

%%% FIGURE 6 %%%
\begin{figure*}
	\includegraphics[width=\linewidth]{H2O}
	\caption{Error with respect to FCI for the singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{H2O} at equilibrium geometry computed at various levels of theory with the aug-cc-pVDZ basis.
	The geometry and the reference FCI values have been extracted from the QUEST database. \cite{Veril_2021}
	Raw data are reported in {\SupInf}.}
	\label{fig:H2O}
\end{figure*}
%%% %%% %%% %%%

As a third and final example, we compute the excitation energies associated with the two lowest singlet and two lowest triplet excited states of water at equilibrium geometry (see Fig.~\ref{fig:H2O}).
Note that all these excited states are of Rydberg nature and correspond to $n \to 3s$ and $n \to 3p$ transitions for the $B_1$ and $A_2$ states, respectively. \cite{Loos_2018a}
In addition to the BSE-based models studied in the present manuscript, we have selected well-known wave function methods, \cite{Bene_1971,Head-Gordon_1994,Head-Gordon_1995,Dreuw_2005} namely CIS, CIS(D), TDHF, and FCI (taken as reference) and computed the excitation energies of these transitions.
It is worth mentioning here that the TDHF (or RPAx) equations within the TDA are strictly equivalent to the CIS equation, \cite{Dreuw_2005} and that CIS(D) is a simple perturbative doubles correction to CIS, and can be considered as an excited-state analog of second-order M{\o}ller-Plesset perturbation theory. \cite{Head-Gordon_1994,Head-Gordon_1995}

Two key observations can be made:
i) BSE@$G_0W_0$ is by far the best performer with a slight overestimation of the order of \SI{0.1}{eV} (as compared to FCI);
ii) BSE@$G_0T_0$ systematically underestimates the excitation energies [similarly to CIS(D)] and outperforms CIS, CIS(D) and TDHF for the singlet states only. 
These general trends are also observed for other systems and they nicely evidence the crucial role of the screening in $GW$, hinting that a screened version of the $T$-matrix formalism as proposed in Ref.~\onlinecite{Romaniello_2012} might be a promising way for improvement.

@@

In order to derive the dynamical kernel $\tilde{T}_{ib,aj}$ given in Eq.~\eqref{eq:dT}, we follow Ref.~\onlinecite{Strinati_1980} (see also Ref.~\onlinecite{Loos_2020e}) and start from the equation for the BSE  amplitude
\begin{equation}
\label{eq:Strinati_1}
	\chi_S(1,1') = \int d3d4d5d6L_0(1,4,1',3)\Xi(3,5,4,6) \chi_S(6,5)
\end{equation}
where $L_0$ is given by Eq.~\eqref{eq:L0} and the $T$-matrix kernel is
\begin{equation}
\label{eq:T-matrix_kernel}
	\Xi(3,5,4,6) =  i\frac{\delta \Sigma(3,4)}{\delta G(6,5)}\approx -T(3,5,4,6)
\end{equation}
Equation (\ref{eq:Strinati_1}) is derived by assuming that i) the (resonant) pole $\omega_S=E_S-E_0>0$ of $L$ is isolated from the other poles (which is usually the case for neutral excitations in finite systems), and  ii) the poles of $L_0$ are different form $\omega_S$ (which is also generally the case).
The so-called $T$-matrix self-energy $\Sigma$ is given by Eq.~\eqref{eq:SigGT} with \cite{Martin_2016,Romaniello_2012}
\begin{equation}
\label{eq:T-matrix}
	T(3,8,4,7) =
	- v(3,8)\delta(3,4)\delta(7,8)
	+ v(3,8)\delta(4,8)\delta(3,7)
	+ i \int d1'd2' v(3,8)G(3,1')G(8,2')T(1',2',4,7)
\end{equation}  
where $v$ is the bare Coulomb operator and we neglect the functional derivative $\delta T/\delta G$ in the kernel $\Xi$. 
The first two terms in the right-hand side of Eq.~\eqref{eq:T-matrix} are the Hartree and exchange contributions to the $T$-matrix, whereas the last term is the correlation contribution.
Making the time dependence of Eq.~\eqref{eq:Strinati_1} explicit and defining $T(3,5,4,6)=- \delta(\tau^+_{35})\delta(\tau^+_{64})\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})$, one gets 
\begin{equation}
\label{eq:DYN_T}
	\chi_S(\bx_1,\bx_{1'},\tau_{11'})e^{-i\omega_S(t_1+t_{1'})/2}
	=-i\int d3d4d5d6 G(\bx_1,\bx_3;\tau_{13})G(\bx_4,\bx_{1'};\tau_{11'})
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}
\end{equation}
[where $\tau_{ij}=t_i-t_j$ and $\tau^+_{ij}=t^+_i-t_j$ with $t^+_i=t_i+\eta$ ($\eta \to 0^+$)] and
\begin{multline}
\label{eq:T_bar}
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34}) 
	= v(\bx_3,\bx_5)\delta(\tau_{34})\delta(\bx_3,\bx_4)\delta(\bx_6,\bx_5)-v(\bx_3,\bx_5)\delta(\tau_{34})	\delta(\bx_3,\bx_6)\delta(\bx_4,\bx_5)
	\\
	+ i \int d7 d8 v(\bx_3,\bx_5)G(\bx_3,\bx_7;\tau_{38})G(\bx_5,\bx_8;\tau^+_{38})\Bar{T}(\bx_7,\bx_8,\bx_4,\bx_6;\tau_{84})
\end{multline}
Using the Fourier transform $G(\tau)=\int \frac{d\omega}{2\pi}G(\omega)e^{-i\omega\tau}$, changing variable from $t_3$ to $\tau_{34}$ and taking the limit $t_{1'}=t_1^+$, we have
\begin{multline}
	\chi_S(\bx_1,\bx_{1'},0^-)
	= -i\int d\bx_3 d\bx_4 d\bx_5 d\bx_6\int d\tau_{34} \int \frac{d\omega'}{2\pi}G(\bx_1,\bx_3;\omega'+\omega_S)G(\bx_4,\bx_{1'};\omega')e^{i\omega'\tau_{34}}
	\\
	 \times \Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{-i\omega_S\tau_{34}/2}
\end{multline}
Using the Lehman representation of the one-body Green's function in the quasiparticle approximation given by Eq.~\eqref{eq:G}, and multiply the left- and right-hand sides by $(\e{a}{}-\e{i}{}-\omega_S)\int d\bx_1 d\bx^{\prime}_1 \SO{a}(\bx_1)\SO{i}(\bx_1')$, we obtain  
\begin{multline}
\label{eq:General_DYN}
	\qty(\e{a}{}-\e{i}{}-\omega_S)\int d\bx_1d\bx^{\prime}_1 \SO{a}(\bx_1)\SO{i}(\bx_1')  \chi_S(\bx_1,\bx_{1'},0^-) =
	-\int d\bx_3 d\bx_4 d\bx_5 d\bx_6\int d\tau_{34}\SO{a}(\bx_3)\SO{i}(\bx_4)
	\\
	\times \qty[\Theta(\tau_{34})e^{i\e{i}{}\tau_{34}}+\Theta(-\tau_{34})e^{i(\e{a}{}-\omega_S)\tau_{34}} ]
	\Bar{T}(\bx_3,\bx_5,\bx_4,\bx_6;\tau_{34})\chi_S(\bx_6,\bx_5;-\tau_{34})e^{i\omega_S\tau_{34}/2}
\end{multline}
(where $\Theta$ is the Heaviside step function) using the fact that
\begin{equation}
	\Theta(\pm\tau)e^{-i\alpha \tau}=\mp\frac{1}{2\pi i}\lim_{\eta\to 0^+}\int d\omega \frac{1}{\omega-\alpha\pm i\eta}e^{-i\omega\tau}
\end{equation}

@@

\begin{equation}
	\label{eq:egval}
	\qty(\e{a}{}-\e{i}{}-\omega_S) X_{ia,S}+\sum_{jb}X_{jb,S}\mel{ib}{}{aj} - \sum_{jb}X_{jb,S}\widetilde{T}_{ib,aj}(\omega_S)=0
\end{equation}
where the spectral representation of the dynamical $T$-matrix is
\begin{equation}
	\widetilde{T}_{ib,aj}(\omega_S)
	= \frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\Bar{T}^c_{ib,aj}(\omega)e^{-i\omega\eta} \qty[\frac{1}{\omega_S-\omega+\e{j}{}+\e{i}{}+i\eta}+\frac{1}{\omega_S+\omega-\e{b}{}-\e{a}{}+i\eta}] 
 \label{eq:H2preso_eigen_2_T}
\end{equation}
Equation \eqref{eq:egval} represents a non-linear eigenvalue equation to calculate the positive excitation energies of a system, which can be rewritten as
\begin{equation}
\label{eq:H2preso_eigen}
	\sum_{jb} A_{ia,jb}(\omega_S) X_{jb,S}=\omega_S X_{ia,S}
\end{equation}
with
\begin{equation}
	A_{ia,jb}(\omega_S)=(\e{a}{}-\e{i}{})\delta_{ij}\delta_{ab} + \mel{ib}{}{aj} - \widetilde{T}_{ib,aj}(\omega_S).
\end{equation}
If one drops the dynamical part $\widetilde{T}$, one ends up with the usual time-dependent Hartree-Fock (TDHF) equations. \cite{Dreuw_2005}
To calculate the correlation contribution, one must employ Eq.~\eqref{eq:T} in Eq.~\eqref{eq:H2preso_eigen_2_T}, and after integration over the frequency, it yields Eq.~\eqref{eq:dT}.
